RIT Department of Computer Science\\MSc Thesis Pre-Proposal:\\\emph{Implementation of a Push-Based Signal-Function FRP System and Evaluation Interface}

### 2011-08-30 Docs/Preproposal/preproposal.tex
Preproposal draft

\title{RIT Department of Computer Science\\MSc Project/Thesis Pre-Proposal:\\\emph{Implementation of a Push Based Signal Function FRP System and Evaluation Interface}}
\author{Edward Amsden}
\date{\today}

@@

The sections shown below are adapted from the topic analysis forms provided in ``Writing the Doctoral Dissertation" (2nd edition) by Davis and Parker (pages 82-88). Your final document should be 1-2 pages including references. {\bf The final pre-proposal may present the items below in any format, but using prose (not bulleted lists).}

@@

Do a monadic evaluation interface and signal vector typing enable a useful push-based FRP system?

\section{Importance of Research}
Functional reactive programming enables a denotative method of defining reactive systems. The efficient evaluation
of such systems is still an open problem. Signal function style FRP semantics have several advantages over
classic FRP semantics, but current implementations use ``pull'' based evaluation, which has the undesirable consequences of
repeated unnecessary work in evaluating event non-occurrences, forcing input to be synchronous with output, and introducing sampling latency into event occurrences.

A signal function FRP system which uses ``pull'' based evaluation would eliminate these consequences, permitting more
efficient implementation with only the latency of processing, rather than the sampling interval. It would also permit
the decoupling of the input and output of the signal function network.

@@

FRP is a well-explored field. There is a wealth of research on the best semantics and most efficient implementation of FRP systems. I previously surveyed the literature regarding FRP for independent study credit~\cite{Amsden2011}. The proposed research would provide an efficient implementation of an elegant family of FRP semantics, and a compositional and intuitive means of providing input to and using output from such systems.

Signal function FRP or Arrowized FRP (AFRP) is one of two main families of semantics for Functional Reactive 
Programming~\cite{Nilsson2002}. It was recently extended with additional typechecked properties using the notion of signal 
vectors.

@@

I will explore an implementation which permits the separation of evaluation of continuous values and events. The implementation will hold the current signal function network as state in the monadic computation described using the evaluation monad. Event occurrences provided as input or produced within the network will mutate the network state.

The evaluation monad will provide monadic actions for ``pushing'' an event occurrence on an input to the signal function network, stepping time and evaluating continous values, and sampling continuous output values.

@@

I expect to produce an implementation of signal function semantics that permits push-based evaluation. I expect that this implementation will include both constructs for producing a signal function network and a monadic interface for evaluating a signal function network. The constructs will include routing combinators, switching combinators (which will introduce information denoting where mutations to the network by events will take place), and stateful signal functions such as
numeric integration.

It is possible that this interface will also permit multiple inputs to and outputs from a signal function network.

>>>>

\title{RIT Department of Computer Science\\MSc Thesis Pre-Proposal:\\\emph{Implementation of a Push-Based Signal-Function FRP System and Evaluation Interface}}
\author{Edward Amsden}
\date{\today}

@@

@@

I intend to show that a monadic evaluation interface and signal vector typing enable a useful push-based FRP system.

\section{Importance of Research}
Functional reactive programming enables a denotative method of defining reactive systems. The efficient evaluation
of such systems is still an open problem. Signal function style FRP semantics have several advantages over
classic FRP semantics, including the elimination of a large class of space and time leaks, as well as explicit input. However,
current implementations use ``pull'' based evaluation, leading to undesirable consequences:
repeating unnecessary work in evaluating event non-occurrences, forcing input to be synchronous with output, and introducing sampling latency into event occurrences.

A signal function FRP system which uses ``push''-based evaluation would eliminate these negatives, permitting more
efficient implementation with only the latency of processing, rather than the sampling interval. It would also permit
the decoupling of the input and output of the signal function network.

@@

FRP is a well-explored field. There is a wealth of research on the best semantics and most efficient implementation of FRP systems, though almost no discussion or exploration of the evaluation interface (the means of providing input and sampling output). I previously surveyed the literature regarding FRP for independent study credit~\cite{Amsden2011}. The proposed research would provide an efficient implementation of an elegant family of FRP semantics, and a compositional and intuitive means of providing input to and using output from such systems.

Signal function FRP or Arrowized FRP (AFRP) is one of two main families of semantics for Functional Reactive 
Programming~\cite{Nilsson2002}. It was recently extended with additional typechecked properties using the notion of signal 
vectors~\cite{Sculthorpe2011}.

@@

I will explore a semantic model of signal function FRP which considers the signal function network as state and events
as mutators of the network. The network can of course evaluate events as well.

The exploration of this model will lead to the design of an implementation which permits the separation of evaluation of
continuous values and discrete events. The implementation will hold the current signal function network as state in the monadic computation described using the evaluation monad. Event occurrences provided as input or produced within the network will mutate the network state.

The evaluation monad will provide monadic actions for ``pushing'' an event occurrence on an input to the signal function network, stepping time and evaluating continous values, and sampling continuous output values. The evaluation monad is
a key component of the design. It permits much finer control over FRP evaluation that in possible with existing models and
implementations, and enables the push-based evaluation of the signal function network by threading state.

@@

I expect to produce an implementation of signal function semantics that utilizes push-based evaluation. I expect that this implementation will include both constructs for producing a signal function network and a monadic interface for evaluating a signal function network. The constructs will include routing combinators, switching combinators (which will introduce information denoting where mutations to the network by events will take place), and stateful signal functions such as
numeric integration.

I expect to explore type-safe means of permitting multiple inputs and outputs.
